Characterizing subarctic European vegetation for land surface modeling of ozone pollution and climate risk

### 2021-09-27 VegOzone.tex
Add changes to section 5 - not finalized

To estimate potential biases in the assessment of ozone-induced risk on Fennoscandia vegetation, we defined hypothetical temperature response functions representing different levels of assumed acclimation to cold climates. We assume northern Fennoscandia vegetation to be more tolerant towards cold temperatures and less tolerant to heat. Based on probability density functions (PDF) of observed light and temperature, we assume that an acclimation ought to increase the enclosed area between the respective response functions and the PDF. These hypothetical species or PFTs still need verification by experimental data.

In summary, our key findings are:
\begin{enumerate}
   \itemsep0pt
 \item the bespoke parameterizations for subarctic species are important (especially concerning light and temperature and the climate effect on the growing season) to be able to assess \chem{O_3} uptake,
 \item the parameterizations defined in the mapping manual for European regional risk assessment would appear to not adequately capture the physiological responses of subarctic physiology that are important determinants of \chem{[O_3]} sensitivity under prevailing environmental conditions and are likely to substantially underestimate risk,
 \item the ozone reanalysis products tend to underestimate \chem{[O_3]} in Fennoscandia, and
 \item that the climatic conditions promoting ozone damage (particularly extended heatwaves and the timing of the growing season concerning the ozone spring peak) and which are likely to become more frequent under climate change are likely to further enhance the sensitivity of subarctic vegetation to ozone.
\end{enumerate}

@@

Based on the methods described in \citet{ICP:MappingManual2017}, we estimated biomass reductions due to ozone uptake for both 2018 and 2019 and find significant reductions for deciduous trees ($15.5-17.4\,\unit{\%}$) and perennial grassland (above ground, $12.4-13.7\,\unit{\%}$) with respect to estimates based on $\mathrm{POD_1}$ from default parameterizations. For coniferous trees, the results are compatible with the default parameterization estimates and are contained within the range of systematic uncertainties deduced from our bespoke response function ($2.5-3\,\unit{\%}$). These results have to be treated with care because the biomass reduction functions were not established based on subarctic species and might as well be subject to biases.

@@

Cumulative ozone uptake has been reported to reduce stomatal conductance significantly and independently from photosynthesis \citep[e.g.]{BGS:Lombardozzi2012}, with photosynthesis being more strongly affected. This might be due competing effects, as an ozone-induced sluggishness \citep{SR:Hoshika2015} renders stomata unable to fully close and leads to higher stomatal conductance than expected and thus higher ozone uptake as well as transpiration. At present, the MM version of the $\mathrm{DO_3SE}$ model does not account for any ozone-inflicted reduction of stomatal conductance throughout the growing season and is not coupled to photosynthesis. Hence, the presented $\mathrm{POD_y}$ may be regarded as an estimated upper limit.\\

>>>>

To estimate potential biases in the assessment of ozone-induced risk on Fennoscandic vegetation, we defined hypothetical temperature response functions representing different levels of assumed acclimation to cold climates. We assume northern Fennoscandia vegetation to be more tolerant towards cold temperatures and less tolerant to heat. Based on probability density functions (PDF) of observed light and temperature, we assume that an acclimation ought to increase the enclosed area between the respective response functions and the PDF. These hypothetical parameterisations of species or PFTs still need verification by experimental data.

In summary, our key findings are:
\begin{enumerate}
   \itemsep0pt
 \item the bespoke parameterizations for subarctic species are important (especially concerning light and temperature and the climate effect on the growing season) to be able to assess \chem{O_3} uptake,
 \item the parameterizations defined in the mapping manual for European regional risk assessment appear to not adequately capture the physiological responses of subarctic physiology that are important determinants of \chem{[O_3]} sensitivity and are likely to substantially underestimate risk, and
 \item that the climatic conditions promoting ozone damage (particularly extended heatwaves and the timing of the growing season concerning the ozone spring peak) and which are likely to become more frequent under climate change are likely to further enhance the sensitivity of subarctic vegetation to ozone.
\end{enumerate}

@@

Based on the methods described in \citet{ICP:MappingManual2017}, we estimated biomass reductions due to ozone uptake for both 2018 and 2019 and found significant reductions for deciduous trees ($15.5-17.4\,\unit{\%}$) and perennial grassland (above ground, $12.4-13.7\,\unit{\%}$) with respect to estimates based on $\mathrm{POD_1}$ from default parameterizations. For coniferous trees, the adjusted parameterizations are comparable with the default parameterization estimates and are contained within the range of systematic uncertainties deduced from our bespoke response function ($2.5-3\,\unit{\%}$). These results have to be treated with care because the biomass reduction functions were not established based on subarctic species.

@@

Cumulative ozone uptake has been reported to reduce stomatal conductance significantly and independently from photosynthesis \citep[e.g.]{BGS:Lombardozzi2012}, with photosynthesis being more strongly affected. This might be due to competing effects, as an ozone-induced sluggishness \citep{SR:Hoshika2015} renders stomata unable to fully close and leads to higher stomatal conductance than expected and thus higher ozone uptake as well as transpiration. At present, the MM version of the $\mathrm{DO_3SE}$ model does not account for any ozone-inflicted reduction of stomatal conductance throughout the growing season and is not coupled to photosynthesis. Hence, the presented $\mathrm{POD_y}$ may be regarded as an estimated upper limit.\\
